---
title: "BiocJobs: declaring dispatchable jobs inside a Bioconductor package"
author:
- name: Bioconductor Core Team
  email: maintainer@bioconductor.org
package: BiocJobs
abstract: >
  BiocJobs lets a package declare, inside the package itself, the units of
  work it can perform non-interactively — *jobs* — as a YAML specification
  plus a plain R script. From that single declaration BiocJobs runs the job
  locally and generates the artifacts execution infrastructure needs: Galaxy
  tool wrappers, GA4GH TES task templates, Nextflow DSL2 modules, WDL tasks,
  and a machine-readable manifest for registry building. This vignette walks
  a package maintainer through the model end to end, using a toy package
  shipped with BiocJobs so every step here is real and reproducible.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{BiocJobs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# Why declare jobs?

Much of what Bioconductor packages do is interactive and exploratory. Some
of it, though, is *batch-shaped*: a well-defined analysis with file inputs,
file outputs, and a handful of parameters. Today every workflow system that
wants to offer such an analysis
— Galaxy, Nextflow, CWL/WDL engines, cloud batch services — needs a
**hand-written wrapper**, maintained by someone who is usually not the
package author, drifting out of sync with the package at every release.
Excellent hand-written wrappers exist for a few popular packages; the long
tail of Bioconductor will never get that treatment.

BiocJobs inverts the ownership. The **package author** — the person who
knows which entry points make sense non-interactively, what the inputs mean,
and which parameters matter — declares the job once, next to the code,
versioned with the code, tested with the code. Wrappers become build
artifacts rather than maintained code.

A package opts in by adding exactly two kinds of files under
`inst/biocjobs/`:

```
mypackage/
└── inst/
    └── biocjobs/
        ├── my-analysis.yaml      # inputs, outputs, options
        └── scripts/
            └── my-analysis.R    # plain R, pure analysis
```

Nothing else about the package changes: no new imports, no code changes, no
build-system requirements. Packages whose value is interactive exploration
simply do not add the directory — opting out is the default.

Three principles hold the framework together:

1. **The YAML is the single source of truth.** It declares inputs, outputs,
   typed configuration options, resource needs and citations. Build
   infrastructure can enumerate and validate every job in a package tarball
   *without installing or executing anything*.
2. **One runtime contract.** Every input, output and option is passed as a
   `--name value` command-line pair, parsed by `jobParams()` against the
   declaration.
3. **One canonical command.** Every execution target launches the same
   self-locating invocation,
   `Rscript -e 'BiocJobs::execJob("<pkg>", "<job>")' --flag value`, so
   generated artifacts contain no absolute paths and never go stale against
   the installed package version.

This vignette uses `toy`, a miniature package shipped inside BiocJobs purely
to demonstrate and test the framework. It declares one job,
`toy-normalize`, that reads a numeric matrix and writes a normalized one.
Everything shown below actually runs while this vignette is built.

```{r toy-package}
library(BiocJobs)

toy <- system.file("examples", "toy", package = "BiocJobs")
list.files(toy, recursive = TRUE)
```

# The anatomy of a job

## The specification

The specification declares the job's *interface*. It is machine-readable
without evaluating any R code, which is what lets infrastructure enumerate
and validate jobs in a tarball safely.

```{r toy-spec, echo = FALSE, comment = ""}
spec_path <- file.path(toy, "inst", "biocjobs", "toy-normalize.yaml")
cat(readLines(spec_path), sep = "\n")
```

`readJob()` parses one specification, `findJobs()` discovers all of them in
a package. Both accept a package source directory, an installed package
directory, or the name of an installed package.

```{r read-job}
jobs <- findJobs(toy)
names(jobs)

job <- jobs[["toy-normalize"]]
job
```

A `BiocJob` is a plain list, so every declared field is directly
inspectable:

```{r inspect-job}
job$version
vapply(job$options, `[[`, "", "name")
str(job$options[[1]])
unlist(job$resources)
```

The `format` field of an input or output comes from a controlled vocabulary
that maps each format to a Galaxy datatype and a file extension:

```{r formats}
head(jobFormats(), 10)
nrow(jobFormats())
```

Unknown formats are not an error — they pass through verbatim to the
generators — but validation flags them as notes, so typos get caught.

## The script

The script is plain R. Its first line hands control of the interface to the
specification; the rest is analysis code reading `params$<name>`.

```{r toy-script, echo = FALSE, comment = ""}
cat(readLines(jobScript(job)), sep = "\n")
```

Note what is *absent*: no argument parsing, no type checking, no usage
message, no defaults. `jobParams()` supplies all of it from the YAML. The
rules that matter when you write your own:

- **First line is `jobParams()`.** Do not read `commandArgs()` yourself, and
  never re-default a value downstream — the spec's defaults are what
  generated user interfaces already showed the user.
- **Fail loudly, early and specifically.** Non-zero exit is how every engine
  detects failure.
- **Never prompt** and never open an interactive device.
- **Write every declared output**, exactly to `params$<name>`.
- **Log to `stderr()` with `message()`** — engines capture it as the job log.

# The runtime contract

Every input, output and option becomes one `--name value` command-line pair,
and every execution environment launches the same command. `jobCommand()`
builds it; it is the single builder used by the local runner and by the
generators, so all execution paths stay in lockstep.

```{r job-command}
jobCommand(job, params = list(
    matrix = "counts.tsv",
    method = "zscore",
    normalized = "normalized.tsv"
))
```

`execJob()` — the `-e` expression above — resolves the job's script inside
the *installed* host package and sources it. That indirection is why
generated artifacts carry no filesystem paths: the same JSON or XML works on
any machine where the package is installed, and always runs the version of
the script that is installed there.

Inside the script, `jobParams()` parses those arguments *against the
declaration*: it coerces types (booleans arrive as logicals, integers as
integers, floats as doubles), applies defaults, validates choice membership
and numeric bounds, checks that required inputs exist on disk, creates
parent directories of output paths, and rejects unknown flags. Inputs,
outputs and options share one flag namespace.

# Running a job locally

`runJob()` executes a job in a fresh `Rscript` process, straight from a
package source checkout — the host package does not need to be installed,
because the specification path is handed to the child process through the
`BIOCJOBS_SPEC` environment variable. The contract is identical to
production.

First, some input data, written — like everything else in this vignette —
into a temporary directory:

```{r make-input}
wd <- file.path(tempdir(), "biocjobs-vignette")
dir.create(wd, showWarnings = FALSE)

set.seed(1)
counts <- matrix(
    rpois(12, 100), nrow = 4,
    dimnames = list(paste0("gene", 1:4), paste0("sample", 1:3))
)
counts_file <- file.path(wd, "counts.tsv")
write.table(data.frame(id = rownames(counts), counts), counts_file,
            sep = "\t", quote = FALSE, row.names = FALSE)

read.delim(counts_file)
```

Now run the job. Parameters are given by their declared names; anything
omitted falls back to the default in the specification (here, `center` and
`pseudocount`).

```{r run-job}
result <- runJob(
    job,
    params = list(matrix = counts_file,
                  normalized = "normalized.tsv",
                  method = "log2"),
    workdir = wd,
    rscript = file.path(R.home("bin"), "Rscript")
)

result$status
result$outputs
```

`runJob()` returns the exit status, the working directory, and the absolute
path of every declared output; it warns when a declared output was not
produced (engines treat that as job failure). The result:

```{r run-output}
read.delim(result$outputs[["normalized"]])
```

Test the failure paths too — a missing required flag, an out-of-range value,
a malformed input file. Each should produce one clear error line, because
that line is what a Galaxy or Nextflow user will see in their job log.

# Validating specifications

`validateJob()` checks a specification for structural problems and returns a
list of issues; an empty list means the specification is valid. **Errors**
make the job unusable, **notes** are advisory.

```{r validate-ok}
issues <- validateJob(job)
length(issues)          # 0: no errors, no notes
```

`readJob()` validates by default and refuses to return an invalid
specification, so most of the time errors surface as soon as you read a
spec. To inspect a broken specification instead of erroring on it, read it
with `validate = FALSE` and validate explicitly. The specification below has
two real errors (a default outside the declared choices, and a name reused
across sections) plus a handful of notes:

```{r validate-broken}
broken_path <- file.path(wd, "broken-job.yaml")
writeLines(c(
    'biocjobs: "1.0"',
    'name: broken-job',
    'package: toy',
    'title: A specification with mistakes',
    'script: scripts/toy-normalize.R',
    'inputs:',
    '  - name: matrix',
    '    format: tabular',
    'outputs:',
    '  - name: matrix',
    '    format: tsv',
    'options:',
    '  - name: method',
    '    type: choice',
    '    choices: [log2, zscore]',
    '    default: sqrt'
), broken_path)

## the spec's script: path is resolved relative to the YAML, so point it at
## a real script for this demonstration
invisible(file.copy(file.path(toy, "inst", "biocjobs", "scripts"), wd,
                    recursive = TRUE))

issues <- validateJob(readJob(broken_path, validate = FALSE))
for (issue in issues)
    cat(sprintf("%-6s %s\n", issue$severity, issue$message))
```

From a shell — the form to wire into continuous integration next to
`R CMD check` — the same check is one command, and it exits non-zero when
any specification has errors:

```sh
Rscript -e 'BiocJobs::biocjobsCLI()' validate /path/to/mypackage
```

Every other operation in this vignette has a shell equivalent too, so build
infrastructure needs nothing but `Rscript`:

```sh
Rscript -e 'BiocJobs::biocjobsCLI()' list     /path/to/pkg
Rscript -e 'BiocJobs::biocjobsCLI()' run <pkg> my-analysis --counts c.tsv
Rscript -e 'BiocJobs::biocjobsCLI()' galaxy   <pkg> my-analysis --out tool.xml
Rscript -e 'BiocJobs::biocjobsCLI()' tes      <pkg> my-analysis --out task.json
Rscript -e 'BiocJobs::biocjobsCLI()' nextflow <pkg> my-analysis --out module.nf
Rscript -e 'BiocJobs::biocjobsCLI()' wdl      <pkg> my-analysis --out task.wdl
Rscript -e 'BiocJobs::biocjobsCLI()' manifest /path/to/pkg --out manifest.json
```

(`biocjobsCLI()` exits the R session with the command's status when running
non-interactively, which is exactly what a CI step wants — and the reason
the chunks in this vignette call the R functions directly instead.)

# The generators

Each generator turns one declaration into one deterministic artifact.
Commit them or regenerate them at release time, but never hand-edit them:
they carry a generated-by header for a reason.

## Galaxy

`galaxyTool()` returns an `xml2` document; `writeGalaxyTool()` writes it and,
when given the job, stages the files referenced by the spec's `tests:` into
a `test-data/` directory beside the XML — the layout `planemo test` expects.

```{r galaxy}
xml_file <- file.path(wd, "toy_normalize.xml")
writeGalaxyTool(galaxyTool(job), xml_file, job = job)

length(readLines(xml_file))                       # the whole tool
cat(head(readLines(xml_file), 28), sep = "\n")    # its first 28 lines
```

The generated tool is complete: typed parameters (`data`, `select`,
`boolean`, `integer`/`float` with bounds, `text` with sanitizers and
validators), outputs with Galaxy datatypes and labels, `<tests>` from the
specification's test cases, DOI citations, a `bioconductor` xref, and
a `<container type="docker">` requirement naming the job's image
that Galaxy resolves to conda environments or mulled BioContainers images.

The tool version follows the IUC convention with the wrapped package version
leading, so regenerating after a Bioconductor release always yields a new
tool version:

```{r galaxy-version}
xml2::xml_attr(xml2::xml_root(galaxyTool(job)), "version")
```

One subtlety worth knowing: Galaxy's default text sanitizer strips
characters such as `~`, which would silently corrupt a design formula like
`~ condition` into `condition`. Declaring `allow_chars: ["~"]` on a string
option makes the generated wrapper extend the sanitizer.

## GA4GH TES

`tesTask()` builds a [GA4GH TES][tes]
v1.1 task, ready to `POST` to a `/tasks` endpoint (Funnel, TESK, cloud
implementations). Declared inputs and outputs become `tesInput`/`tesOutput`
entries staged at fixed container paths, `resources:` becomes
`tesResources`, and the final executor runs the canonical command inside a
Bioconductor container.

```{r tes}
task <- tesTask(
    job,
    inputs  = c(matrix = "s3://bucket/counts.tsv"),
    outputs = c(normalized = "s3://bucket/normalized.tsv")
)
names(task)

json_file <- file.path(wd, "toy-normalize.tes.json")
writeTesTask(task, json_file)
cat(head(readLines(json_file), 25), sep = "\n")
```

Because the generic `bioconductor/bioconductor_docker` image ships no
analysis packages, a *bootstrap* executor is prepended that installs the
host package, BiocJobs and any declared `depends` with `BiocManager`, so the
task is genuinely executable out of the box. Point `container:` (or the
`image` argument) at a purpose-built image and the bootstrap disappears:

```{r tes-executors}
length(task$executors)
unlist(task$executors[[length(task$executors)]]$command)

img <- "quay.io/biocontainers/bioconductor-toy:1.2.3"
length(tesTask(job, image = img)$executors)
```

Values that are not known at generation time are emitted as
`{{...}}` placeholders and the task is tagged `biocjobs.template`, so the
JSON acts as a *submission template*: fill in the URLs (and any unset
required options) and submit it.

```{r tes-template}
template <- tesTask(job)
template$inputs[[1]]$url
unlist(template$tags)
```

If an unfilled placeholder ever reaches a running job, `jobParams()` rejects
it by name rather than letting a literal `{{...}}` leak into the analysis.

## Nextflow

`nextflowModule()` returns a DSL2 module as text, or writes it when given a
`file`: one `process`, file inputs as `path`, options as `val` (annotated
with their types and the specification's defaults), outputs emitted under
stable names, resource directives, and a `stub:` block so pipelines can be
smoke-tested with `-stub-run` before touching real data or containers.

The whole artifact is short enough to show in full:

```{r nextflow}
module <- nextflowModule(job)
cat(strsplit(module, "\n")[[1]], sep = "\n")
```

Two things to know when calling the module: process inputs are
**positional** (Nextflow has no named process-input syntax), so the calling
workflow passes the file inputs and then every option in the order they
appear in the specification; and each option is a required `val` with no
in-module default, its spec default shown in a comment. Reordering options
in a future spec version therefore changes the call signature — bump the
job's `version:` when you do.

Following the nf-core convention, the file inputs arrive together in a
single `tuple val(meta), path(...)` input led by a `meta` map, the outputs
carry that map back out, and the process `tag` is `${meta.id}` so Nextflow's
progress output names the unit of work rather than the process:

```
include { TOY_NORMALIZE } from './modules/toy_normalize.nf'

workflow {
    ch = channel.of([ [id: 'sample1'], file(params.matrix) ])
    TOY_NORMALIZE(ch, 'log2', false, 1.0)
    TOY_NORMALIZE.out.normalized.view()   // [ [id:sample1], normalized.tsv ]
}
```

Pass `meta = FALSE` to `nextflowModule()` for pipelines that do not use meta
maps; the module then takes plain `path` inputs and tags with the first
input file's name.

## WDL

`wdlTask()` generates a [WDL](https://openwdl.org) 1.0 task — the version
with the widest engine support (Cromwell, miniwdl, Terra, dxWDL). Spec
defaults become WDL defaults, so options *without* a default become required
WDL inputs enforced by the engine itself; `parameter_meta` carries the
labels, help, choices and bounds that the WDL type system cannot express.

```{r wdl}
wdl <- wdlTask(job)
cat(head(strsplit(wdl, "\n")[[1]], 27), sep = "\n")
```

The rest of the task carries the `runtime` block built from `resources:`,
the `parameter_meta` entries, and a `meta` block naming the package, job and
job version:

```{r wdl-meta}
cat(tail(strsplit(wdl, "\n")[[1]], 22), sep = "\n")
```

```
import "toy_normalize.wdl" as jobs

workflow normalize {
    call jobs.toy_normalize { input: matrix = matrix_file, method = "log2" }
}
```

## HTCondor

`htcondorSubmit()` generates an [HTCondor](https://htcondor.org) submit
description together with the executable script it runs. Declared inputs
become `transfer_input_files`, outputs `transfer_output_files`, `resources:`
the `request_*` knobs, and the container an image under the container
universe.

```{r htcondor}
sub <- htcondorSubmit(job, image = "bioconductor/bioconductor_docker:devel")
cat(sub)
```

HTCondor has no typed parameter surface, so this is a concrete submission
rather than a typed template like the Galaxy tool or the WDL task. Option
values are written into the script — the declared default where there is
one, and otherwise the same `{{options.<name>}}` placeholder `tesTask()`
uses, which `jobParams()` rejects by name if it ever reaches a running job.
Note that HTCondor transfers inputs into the job's scratch directory by
basename, so the script refers to every file by basename:

```{r htcondor-script}
condor_dir <- file.path(wd, "condor")
dir.create(condor_dir, showWarnings = FALSE)
invisible(htcondorSubmit(job, file = file.path(condor_dir,
                                               "toy-normalize.sub")))
list.files(condor_dir)
cat(readLines(file.path(condor_dir, "toy-normalize.sh")), sep = "\n")
```

Submit it directly with `condor_submit toy-normalize.sub`, or hand the pair
to the [submitr](https://cran.r-project.org/package=submitr) package, which
stages and submits to an HTC submit node such as CHTC:

```
cfg <- submitr::htc_config()
submitr::htc_upload(files = c("toy-normalize.sub", "toy-normalize.sh"),
                    config = cfg)
submitr::htc_submit(submit_file = "toy-normalize.sub", config = cfg)
```

# The manifest and the registry story

`jobManifest()` summarizes every job a package declares — the full typed
interface, resources, container, and canonical commands — as one JSON
document per package.

```{r manifest}
manifest <- jobManifest(toy)
manifest$package
manifest$version
lengths(manifest$jobs[[1]][c("inputs", "outputs", "options")])
manifest$jobs[[1]]$command
```

```{r manifest-json}
manifest_file <- file.path(wd, "manifest.json")
invisible(jobManifest(toy, file = manifest_file))
cat(head(readLines(manifest_file), 22), sep = "\n")
```

The manifest is the **aggregation unit**. Because discovery needs no code
evaluation, build infrastructure can collect manifests across every package
in a release — straight from tarballs — into a Bioconductor-wide registry of
dispatchable jobs, from which TES templates and Galaxy tool sheds regenerate
automatically at every release:

```
package tarballs ──▶ findJobs()/validateJob() ──▶ per-package manifests
   ──▶ registry ──▶ {TES, Galaxy, Nextflow, WDL} per release
```

# Adding a job to your own package

`jobSkeleton()` scaffolds the two files in a package source tree. Both are
working templates: the YAML demonstrates every specification field with
comments, and the script demonstrates the contract. Existing files are never
overwritten unless you pass `overwrite = TRUE`.

```{r skeleton}
pkg_dir <- file.path(tempdir(), "mypackage")
dir.create(pkg_dir, showWarnings = FALSE)
writeLines(c("Package: mypackage", "Version: 0.0.1"),
           file.path(pkg_dir, "DESCRIPTION"))

jobSkeleton("my-analysis", pkg = pkg_dir)
list.files(pkg_dir, recursive = TRUE)
```

```{r skeleton-script, comment = ""}
cat(readLines(file.path(pkg_dir, "inst", "biocjobs", "scripts",
                        "my-analysis.R")), sep = "\n")
```

From there the loop is: edit the YAML, write the analysis, `runJob()` it on
small committed test data, `validateJob()` it, declare `tests:` for the
Galaxy wrapper, and regenerate the artifacts. The
[developer guide][guide]
documents every specification field, the release checklist, and a
troubleshooting table.

# A command line for free (optional)

The same declarations double as a human-facing command line through
[BiocExecute](https://github.com/BiocCodingCollaborations/BiocExecute), a
community CLI framework built on
[Rapp](https://cran.r-project.org/package=Rapp). It compiles every job a
package declares into a subcommand of one launcher, with `--help` text,
option types and defaults all drawn from the YAML — **you write nothing
extra**.

The chunks in this section are shown but not evaluated: BiocExecute is not a
Bioconductor package (it lives on GitHub, and the BiocJobs integration is on
its `feat/biocjobs-specs` branch), so it cannot be a dependency of this
vignette.

```{r biocexecute-compile, eval = FALSE}
## in your package source dir: reads inst/biocjobs/,
## writes exec/<Package>.R
BiocExecute::execCompile()

## once, after the package is installed: puts the launcher on the user's PATH
BiocExecute::execInstall("mypackage")
```

```sh
$ mypackage --help
Usage: mypackage <COMMAND>
Commands:
  my-analysis  One-line title from the specification

$ mypackage my-analysis --input1 data.tsv --method fast
```

The division of labour is deliberate: **BiocJobs owns the declaration and
its validation**, **BiocExecute/Rapp own the shell ergonomics** (subcommand
dispatch, `--help`, PATH launchers). At run time the generated subcommand
hands its parsed values straight to `execJob()`:

```{r biocexecute-values, eval = FALSE}
BiocJobs::execJob("mypackage", "my-analysis",
                  values = list(input1 = "data.tsv", method = "fast"))
```

so required options, choice membership and file checks are enforced by the
same specification that drives every workflow target — one contract,
identical error messages. A job name becomes a subcommand by mapping `-` to
`_` internally and back for display; `cliJobName()` reports the token, and
`validateJob()` emits a note for the rare name that cannot survive the round
trip.

```{r cli-name}
cliJobName(job)
```

Maintainers who want the CLI run `execCompile()` once and commit the
generated `exec/<Package>.R`; maintainers who do not, ignore it — the
workflow targets never depend on it.

# Containers and dependencies

At run time a job needs three things in its environment: R, the host package
(plus anything listed in `depends:`), and BiocJobs — whose only runtime role
is `jobParams()`, and which brings just **yaml**, **jsonlite** and **xml2**
along. The targets get there differently:

- **TES, Nextflow, WDL** default to
  `bioconductor/bioconductor_docker:<current release>`, with the TES
  bootstrap installing what is missing at task start. For production, point
  `container:` at a purpose-built image — for Bioconductor packages the
  auto-built `quay.io/biocontainers/bioconductor-<pkg>` images are a natural
  base — and the bootstrap disappears.
- **Galaxy** declares a `<container type="docker">` requirement naming the
  same image, so the tool runs in the same environment as every other
  target. Bioconda package requirements are not emitted for now: they would
  describe a second, separately-resolved environment, and BiocJobs is not
  yet on bioconda.

```{r container}
job$container            # NULL: the job accepts the default image
manifest$jobs[[1]]$container
```

# Scope: when *not* to declare a job

Declare a job when the analysis is batch-shaped: inputs are files (or can
reasonably be serialized to files), outputs are files, every decision the
user makes can be expressed as a typed option up front, and the code path
runs unattended from start to finish. Differential expression, normalization,
peak calling, amplicon denoising, quantification import, batch correction and
annotation all fit.

Do **not** declare jobs for work whose value is interactive: visualization
browsers, iterative model tuning with a human in the loop, anything needing a
live R session mid-analysis. If some computation *inside* an interactive flow
is batch-shaped — fitting a model, scoring a matrix — consider exposing just
that.

One package can declare several jobs (one YAML plus script pair each, sharing
one name namespace), and a job need not expose everything a function can do:
expose the parameters that matter for batch use, hard-code sensible choices
for the rest, and keep the full flexibility in your R API.

Version 1.0 of the specification deliberately keeps the model small:
single-file inputs and outputs, five option types, one executor per job.
Multi-file inputs and Galaxy collections, a reserved `threads` option wired
to `resources.cpus`, and further targets such as a CWL `CommandLineTool` are
on the roadmap — the Nextflow and WDL generators demonstrate how little a new
target costs, because they consume the same specifications.

# Where to go next

- The
  [developer guide][guide]
  walks a maintainer from an empty directory to generated, validated
  wrappers, with the complete field reference, a release checklist, a
  troubleshooting table and an FAQ.
- The **DESeq2 worked example** in the BiocJobs repository
  ([`examples/DESeq2/`][deseq2])
  is the realistic counterpart to the toy job used here: 2 inputs, 3 outputs
  and 9 typed options exercising every option type, a defensively written
  analysis script, simulated test data, and all the generated artifacts under
  `generated/` — a Galaxy tool validated against Galaxy's official XSD, a TES
  task validated against the GA4GH TES 1.1 schema, a Nextflow module that
  passes `nextflow lint` and runs under `-stub-run`, and a WDL task that
  passes `miniwdl check`. It is not installed with the package (and needs
  DESeq2), which is why this vignette uses the toy job instead.
- `help(package = "BiocJobs")` for the reference documentation of all
  exported functions.

# Documentation status

This documentation was drafted with AI assistance and has not yet been fully
reviewed. It may be inaccurate or out of date in places until every claim has
been checked against the implementation. Treat it as provisional while this
notice is present; it will be removed once the review is complete.

# Session information

```{r session-info}
sessionInfo()
```

[guide]: https://github.com/almahmoud/BiocJobs/blob/main/docs/developer-guide.md
[deseq2]: https://github.com/almahmoud/BiocJobs/tree/main/examples/DESeq2
[tes]: https://github.com/ga4gh/task-execution-schemas
