---
title: "BiocJobs: declaring dispatchable jobs inside a Bioconductor package"
author:
- name: Alexandru Mahmoud
  email: almahmoudresearch@gmail.com
date: "`r format(Sys.Date(), '%d %B %Y')`"
package: BiocJobs
abstract: >
  BiocJobs lets a package declare, inside the package itself, the units of
  work it can perform non-interactively (*jobs*) as a YAML specification
  plus a plain R script. From that single declaration BiocJobs runs the job
  locally and generates the artifacts execution infrastructure needs: Galaxy
  tool wrappers, GA4GH TES tasks, Nextflow DSL2 modules, WDL tasks, HTCondor
  submit files, Kubernetes Jobs and a machine-readable manifest for registry
  building. This vignette walks a package maintainer through the model end
  to end, using a toy package shipped with BiocJobs so every step here is
  real and reproducible.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Declaring dispatchable jobs with BiocJobs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# Why declare jobs?

Batch-shaped analyses (files in, files out, a few parameters) reach Galaxy,
Nextflow and other workflow systems through hand-written wrappers maintained
outside the package. With BiocJobs the package author declares the job once,
next to the code, and the wrappers are generated from that declaration.

A package opts in by adding exactly two kinds of files under
`inst/biocjobs/`:

```
mypackage/
└── inst/
    └── biocjobs/
        ├── my-analysis.yaml      # inputs, outputs, options
        └── scripts/
            └── my-analysis.R    # plain R, pure analysis
```

Nothing else about the package changes: no new imports, no code changes, no
build-system requirements. Packages whose value is interactive exploration
do not add the directory.

Three principles hold the framework together:

1. **The YAML is the single source of truth.** It declares inputs, outputs,
   typed configuration options, resource needs and citations. Build
   infrastructure can enumerate and validate every job in a package tarball
   *without installing or executing anything*.
2. **One runtime contract.** Every input, output and option is passed as a
   `--name value` command-line pair, parsed by `jobParams()` against the
   declaration.
3. **One canonical command.** Every execution target launches the same
   self-locating invocation,
   `Rscript -e 'BiocJobs::execJob("<pkg>", "<job>")' --flag value`, so
   generated artifacts contain no absolute paths and never go stale against
   the installed package version.

BiocJobs is built for Bioconductor packages. Generated artifacts default to
the `bioconductor/bioconductor_docker` image of the release that
`BiocManager::version()` reports (except the Kubernetes Job, which needs an
image), the TES bootstrap installs the host
package and its declared dependencies with `BiocManager`, and discovery and
validation read only the YAML, so the Bioconductor build system can check
every package's jobs without installing them.

# Installation

```{r install, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("BiocJobs")
```

This vignette uses `toy`, a miniature package shipped inside BiocJobs purely
to demonstrate and test the framework. It declares one job,
`toy-normalize`, that reads a numeric matrix and writes a normalized one.
The chunks below run while this vignette is built, except the one in the
section on command-line front ends.

```{r toy-package}
library(BiocJobs)

toy <- system.file("examples", "toy", package = "BiocJobs")
list.files(toy, recursive = TRUE)
```

# The anatomy of a job

## The specification

The specification declares the job's *interface*. It is machine-readable
without evaluating any R code, which is what lets infrastructure enumerate
and validate jobs in a tarball safely.

```{r toy-spec, echo = FALSE, comment = ""}
spec_path <- file.path(toy, "inst", "biocjobs", "toy-normalize.yaml")
cat(readLines(spec_path), sep = "\n")
```

`readJob()` parses one specification, `findJobs()` discovers all of them in
a package. Both accept a package source directory, an installed package
directory, or the name of an installed package.

```{r read-job}
jobs <- findJobs(toy)
names(jobs)

job <- jobs[["toy-normalize"]]
job
```

A `BiocJob` is a plain list, so every declared field is directly
inspectable:

```{r inspect-job}
job$version
vapply(job$options, `[[`, "", "name")
str(job$options[[1]])
unlist(job$resources)
```

The `format` field of an input or output comes from a controlled vocabulary
that maps each format to a Galaxy datatype and a file extension:

```{r formats}
head(jobFormats(), 10)
nrow(jobFormats())
```

Unknown formats are not an error (they pass through verbatim to the
generators), but validation flags them as notes, so typos get caught.

## The script

The script is plain R. Its first line hands control of the interface to the
specification; the rest is analysis code reading `params$<name>`.

```{r toy-script, echo = FALSE, comment = ""}
cat(readLines(jobScript(job)), sep = "\n")
```

Note what is *absent*: no argument parsing, no type checking, no usage
message, no defaults. `jobParams()` supplies all of it from the YAML. The
rules that matter when you write your own:

- **First line is `jobParams()`.** Do not read `commandArgs()` yourself, and
  never re-default a value downstream: the spec's defaults are what
  generated user interfaces already showed the user.
- **Fail loudly, early and specifically.** Non-zero exit is how every engine
  detects failure.
- **Never prompt** and never open an interactive device.
- **Write every declared output**, exactly to `params$<name>`.
- **Log to `stderr()` with `message()`**; engines capture it as the job log.

# The runtime contract

Every input, output and option becomes one `--name value` command-line pair,
and every generated artifact launches the same command. `jobCommand()`
builds the command line that `tesTask()`, `kubernetesJob()` and
`jobManifest()` embed:

```{r job-command}
jobCommand(job, params = list(
    matrix = "counts.tsv",
    method = "zscore",
    normalized = "normalized.tsv"
))
```

`execJob()`, the `-e` expression above, resolves the job's script inside
the *installed* host package and sources it. That indirection is why
generated artifacts carry no filesystem paths: the same JSON or XML works on
any machine where the package is installed, and always runs the version of
the script that is installed there.

Inside the script, `jobParams()` parses those arguments *against the
declaration*: it coerces types (booleans arrive as logicals, integers as
integers, floats as doubles), applies defaults, validates choice membership
and numeric bounds, checks that required inputs exist on disk, creates
parent directories of output paths, and rejects unknown flags. Inputs,
outputs and options share one flag namespace.

# Running a job locally

`runJob()` executes a job in a fresh `Rscript` process, straight from a
package source checkout. The host package does not need to be installed,
because the specification path is handed to the child process through the
`BIOCJOBS_SPEC` environment variable. The contract is identical to
production.

First, some input data, written (like everything else in this vignette)
into a temporary directory:

```{r make-input}
wd <- file.path(tempdir(), "biocjobs-vignette")
dir.create(wd, showWarnings = FALSE)

set.seed(1)
counts <- matrix(
    rpois(12, 100), nrow = 4,
    dimnames = list(paste0("gene", 1:4), paste0("sample", 1:3))
)
counts_file <- file.path(wd, "counts.tsv")
write.table(data.frame(id = rownames(counts), counts), counts_file,
            sep = "\t", quote = FALSE, row.names = FALSE)

read.delim(counts_file)
```

Now run the job. Parameters are given by their declared names; anything
omitted falls back to the default in the specification (here, `center` and
`pseudocount`).

```{r run-job}
result <- runJob(
    job,
    params = list(matrix = counts_file,
                  normalized = "normalized.tsv",
                  method = "log2"),
    workdir = wd,
    rscript = file.path(R.home("bin"), "Rscript")
)

result$status
result$outputs
```

`runJob()` returns the exit status, the working directory, and the absolute
path of every declared output; it warns when a declared output was not
produced (engines treat that as job failure). The result:

```{r run-output}
read.delim(result$outputs[["normalized"]])
```

Test the failure paths too: a missing required flag, an out-of-range value,
a malformed input file. Each should produce one clear error line, because
that line is what a Galaxy or Nextflow user will see in their job log.

# Validating specifications

`validateJob()` checks a specification for structural problems and returns a
list of issues; an empty list means the specification is valid. **Errors**
make the job unusable, **notes** are advisory.

```{r validate-ok}
issues <- validateJob(job)
length(issues)          # 0: no errors, no notes
```

`readJob()` validates by default and refuses to return an invalid
specification, so most of the time errors surface as soon as you read a
spec. To inspect a broken specification instead of erroring on it, read it
with `validate = FALSE` and validate explicitly. The specification below has
two real errors (a default outside the declared choices, and a name reused
across sections) plus a handful of notes:

```{r validate-broken}
broken_path <- file.path(wd, "broken-job.yaml")
writeLines(c(
    'biocjobs: "1.0"',
    'name: broken-job',
    'package: toy',
    'title: A specification with mistakes',
    'script: scripts/toy-normalize.R',
    'inputs:',
    '  - name: matrix',
    '    format: tabular',
    'outputs:',
    '  - name: matrix',
    '    format: tsv',
    'options:',
    '  - name: method',
    '    type: choice',
    '    choices: [log2, zscore]',
    '    default: sqrt'
), broken_path)

## the spec's script: path is resolved relative to the YAML, so point it at
## a real script for this demonstration
invisible(file.copy(file.path(toy, "inst", "biocjobs", "scripts"), wd,
                    recursive = TRUE))

issues <- validateJob(readJob(broken_path, validate = FALSE))
for (issue in issues)
    cat(sprintf("%-6s %s\n", issue$severity, issue$message))
```

From a shell, the form to wire into continuous integration next to
`R CMD check`, the same check is one command, and it exits non-zero when
any specification has errors:

```sh
Rscript -e 'BiocJobs::biocjobsCLI()' validate /path/to/mypackage
```

Every other operation in this vignette has a shell equivalent too, so build
infrastructure needs nothing but `Rscript`:

```sh
Rscript -e 'BiocJobs::biocjobsCLI()' list     /path/to/pkg
Rscript -e 'BiocJobs::biocjobsCLI()' run <pkg> my-analysis --counts c.tsv
Rscript -e 'BiocJobs::biocjobsCLI()' galaxy   <pkg> my-analysis --out tool.xml
Rscript -e 'BiocJobs::biocjobsCLI()' tes      <pkg> my-analysis --out task.json
Rscript -e 'BiocJobs::biocjobsCLI()' nextflow <pkg> my-analysis --out module.nf
Rscript -e 'BiocJobs::biocjobsCLI()' wdl      <pkg> my-analysis --out task.wdl
Rscript -e 'BiocJobs::biocjobsCLI()' manifest /path/to/pkg --out manifest.json
```

`biocjobsCLI()` quits R with the command's exit status when run
non-interactively, so the chunks here call the R functions directly.

# The generators

Each generator turns one declaration into one deterministic artifact.
Commit them or regenerate them at release time, but never hand-edit them.

## Galaxy

`galaxyTool()` returns an `xml2` document; `writeGalaxyTool()` writes it and,
when given the job, stages the files referenced by the spec's `tests:` into
a `test-data/` directory beside the XML, the layout `planemo test` expects.

```{r galaxy}
xml_file <- file.path(wd, "toy_normalize.xml")
writeGalaxyTool(galaxyTool(job), xml_file, job = job)

length(readLines(xml_file))                       # the whole tool
cat(head(readLines(xml_file), 28), sep = "\n")    # its first 28 lines
```

The generated tool is complete: typed parameters (`data`, `select`,
`boolean`, `integer`/`float` with bounds, `text` with sanitizers and
validators), outputs with Galaxy datatypes and labels, `<tests>` from the
specification's test cases, DOI citations, a `bioconductor` xref, and
a `<container type="docker">` requirement naming the job's image.

The tool version follows the IUC convention with the wrapped package version
leading, so regenerating after a Bioconductor release always yields a new
tool version:

```{r galaxy-version}
xml2::xml_attr(xml2::xml_root(galaxyTool(job)), "version")
```

Galaxy's default text sanitizer strips
characters such as `~`, which would silently corrupt a design formula like
`~ condition` into `condition`. Declaring `allow_chars: ["~"]` on a string
option makes the generated wrapper extend the sanitizer.

## GA4GH TES

`tesTask()` builds a [GA4GH TES][tes]
v1.1 task, ready to `POST` to a `/tasks` endpoint (Funnel, TESK, cloud
implementations). Declared inputs and outputs become `tesInput`/`tesOutput`
entries staged at fixed container paths, `resources:` becomes
`tesResources`, and the final executor runs the canonical command inside a
Bioconductor container.

```{r tes}
task <- tesTask(
    job,
    inputs  = c(matrix = "s3://bucket/counts.tsv"),
    outputs = c(normalized = "s3://bucket/normalized.tsv")
)
names(task)

json_file <- file.path(wd, "toy-normalize.tes.json")
writeTesTask(task, json_file)
cat(head(readLines(json_file), 25), sep = "\n")
```

On the default `bioconductor/bioconductor_docker` image the task starts with
a bootstrap executor that installs the host package, BiocJobs and any
declared `depends` with `BiocManager`. Setting `container:` or `image`
removes it:

```{r tes-executors}
length(task$executors)
unlist(task$executors[[length(task$executors)]]$command)

img <- "ghcr.io/example/toy:1.0"
length(tesTask(job, image = img)$executors)
```

Values that are not known at generation time are emitted as
`{{...}}` placeholders and the task is tagged `biocjobs.template`, so the
JSON acts as a *submission template*: fill in the URLs (and any unset
required options) and submit it.

```{r tes-template}
template <- tesTask(job)
template$inputs[[1]]$url
unlist(template$tags)
```

If an unfilled placeholder ever reaches a running job, `jobParams()` rejects
it by name rather than letting a literal `{{...}}` leak into the analysis.

## Nextflow

`nextflowModule()` returns a DSL2 module with one `process`: the file inputs
arrive in one `tuple val(meta), path(...)` input, options are `val` inputs
annotated with their types and defaults, outputs are emitted under their
declared names together with `meta`, resources become directives, and a
`stub:` block supports `-stub-run`.

The whole artifact is short enough to show in full:

```{r nextflow}
module <- nextflowModule(job)
cat(strsplit(module, "\n")[[1]], sep = "\n")
```

Two things to know when calling the module: process inputs are
**positional** (Nextflow has no named process-input syntax), so the calling
workflow passes the file inputs and then every option in the order they
appear in the specification; and each option is a required `val` with no
in-module default, its spec default shown in a comment. Reordering options
in a future spec version therefore changes the call signature; bump the
job's `version:` when you do. A calling workflow:

```
include { TOY_NORMALIZE } from './modules/toy_normalize.nf'

workflow {
    ch = channel.of([ [id: 'sample1'], file(params.matrix) ])
    TOY_NORMALIZE(ch, 'log2', false, 1.0)
    TOY_NORMALIZE.out.normalized.view()   // [ [id:sample1], normalized.tsv ]
}
```

Pass `meta = FALSE` to `nextflowModule()` for pipelines that do not use meta
maps; the module then takes plain `path` inputs and tags with the first
input file's name.

## WDL

`wdlTask()` generates a [WDL](https://openwdl.org) 1.0 task. Spec
defaults become WDL defaults, so options *without* a default become required
WDL inputs enforced by the engine itself; `parameter_meta` carries the
labels, help, choices and bounds that the WDL type system cannot express.

```{r wdl}
wdl <- wdlTask(job)
cat(head(strsplit(wdl, "\n")[[1]], 27), sep = "\n")
```

The rest of the task carries the `runtime` block built from `resources:`,
the `parameter_meta` entries, and a `meta` block naming the package, job and
job version:

```{r wdl-meta}
cat(tail(strsplit(wdl, "\n")[[1]], 22), sep = "\n")
```

```
import "toy_normalize.wdl" as jobs

workflow normalize {
    input { File matrix_file }
    call jobs.toy_normalize { input: matrix = matrix_file, method = "log2" }
}
```

## HTCondor

`htcondorSubmit()` generates an [HTCondor](https://htcondor.org) submit
description together with the executable script it runs. Declared inputs
become `transfer_input_files`, outputs `transfer_output_files`, `resources:`
the `request_*` knobs, and the container an image under the container
universe.

```{r htcondor}
sub <- htcondorSubmit(job, image = "bioconductor/bioconductor_docker:devel")
cat(sub)
```

HTCondor has no typed parameter surface, so this is a concrete submission
rather than a typed template like the Galaxy tool or the WDL task. Option
values are written into the script: the declared default where there is
one, and otherwise the same `{{options.<name>}}` placeholder `tesTask()`
uses, which `jobParams()` rejects by name if it ever reaches a running job.
HTCondor transfers inputs into the job's scratch directory by
basename, so the script refers to every file by basename:

```{r htcondor-script}
condor_dir <- file.path(wd, "condor")
dir.create(condor_dir, showWarnings = FALSE)
invisible(htcondorSubmit(job, file = file.path(condor_dir,
                                               "toy-normalize.sub")))
list.files(condor_dir)
cat(readLines(file.path(condor_dir, "toy-normalize.sh")), sep = "\n")
```

Submit it directly with `condor_submit toy-normalize.sub`, or hand the pair
to the [submitr](https://cran.r-project.org/package=submitr) package, which
stages and submits to an HTC submit node such as CHTC:

```
cfg <- submitr::htc_config()
submitr::htc_upload(files = c("toy-normalize.sub", "toy-normalize.sh"),
                    config = cfg)
submitr::htc_submit(submit_file = "toy-normalize.sub", config = cfg)
```

## Kubernetes

`kubernetesJob()` generates a Kubernetes `batch/v1` Job. Init containers
download the inputs with `curl` into a shared volume, and the job container
runs the canonical command with inputs under `/biocjob/inputs` and outputs
under `/biocjob/outputs`.

```{r kubernetes}
k8s <- kubernetesJob(job,
                     inputs = c(matrix = "https://example.org/counts.tsv"),
                     image = "ghcr.io/example/toy:1.0")
cat(writeKubernetesJob(k8s))
```

The image must provide `curl`, BiocJobs and the package; submit the Job
with `kubectl create -f`. Outputs are lost when the job ends unless `claim`
names a PersistentVolumeClaim to write them to.

# The manifest

`jobManifest()` summarizes every job a package declares (the typed
interface, resources, container and canonical commands) as one JSON
document per package.

```{r manifest}
manifest <- jobManifest(toy)
manifest$package
manifest$version
lengths(manifest$jobs[[1]][c("inputs", "outputs", "options")])
manifest$jobs[[1]]$command
```

```{r manifest-json}
manifest_file <- file.path(wd, "manifest.json")
invisible(jobManifest(toy, file = manifest_file))
cat(head(readLines(manifest_file), 22), sep = "\n")
```

Discovery needs no code evaluation, so build infrastructure can collect
manifests from package tarballs into a registry of jobs across packages.

# Adding a job to your own package

`jobSkeleton()` scaffolds the two files in a package source tree. Both are
working templates: the YAML is a valid specification with comments on the
common fields, and the script demonstrates the contract. Existing files are
never overwritten unless you pass `overwrite = TRUE`.

```{r skeleton}
pkg_dir <- file.path(tempdir(), "mypackage")
dir.create(pkg_dir, showWarnings = FALSE)
writeLines(c("Package: mypackage", "Version: 0.0.1"),
           file.path(pkg_dir, "DESCRIPTION"))

jobSkeleton("my-analysis", pkg = pkg_dir)
list.files(pkg_dir, recursive = TRUE)
```

```{r skeleton-script, comment = ""}
cat(readLines(file.path(pkg_dir, "inst", "biocjobs", "scripts",
                        "my-analysis.R")), sep = "\n")
```

From there the loop is: edit the YAML, write the analysis, `runJob()` it on
small committed test data, `validateJob()` it, declare `tests:` for the
Galaxy wrapper, and regenerate the artifacts. The
[developer guide][guide]
documents every specification field, the release checklist, and a
troubleshooting table.

# Command-line front ends

A command-line front end that parses the arguments itself, for example a
[Rapp](https://cran.r-project.org/package=Rapp) application, hands the
parsed values to `execJob()`:

```{r cli-values, eval = FALSE}
BiocJobs::execJob("mypackage", "my-analysis",
                  values = list(input1 = "data.tsv", method = "fast"))
```

The script's `jobParams()` call then uses these values instead of the
command line, so required options, choice membership and file checks come
from the same specification as every workflow target. Values of `NULL` or
`NA` count as not supplied.

A front end that exposes each job as a subcommand maps `-` in the job name
to `_` for the R name and back for display. `cliJobName()` returns the
resulting token, `validateJob()` notes a name that cannot be mapped, and
`jobCommand(style = "cli")` builds the command line for such a launcher:

```{r cli-name}
cliJobName(job)
jobCommand(job, list(matrix = "counts.tsv"), style = "cli")
```

# Containers and dependencies

At run time a job needs three things in its environment: R, the host package
(plus anything listed in `depends:`), and BiocJobs, whose only runtime role
is `jobParams()` and which brings only **yaml**, **jsonlite** and **xml2**
along. The targets get there differently:

- Every target except Kubernetes defaults to
  `bioconductor/bioconductor_docker:<current release>`; on that image the
  TES task first installs what is missing. For production, set `container:`
  to an image with R, the package, its `depends` and BiocJobs installed. The
  Kubernetes Job requires such an image.
- **Galaxy** declares a `<container type="docker">` requirement naming the
  same image, so the tool runs in the same environment as every other
  target. Bioconda package requirements are not emitted for now: they would
  describe a second, separately-resolved environment, and BiocJobs is not
  yet on bioconda.

```{r container}
job$container            # NULL: the job accepts the default image
manifest$jobs[[1]]$container
```

# Scope: when *not* to declare a job

Declare a job when the analysis is batch-shaped: inputs are files (or can
reasonably be serialized to files), outputs are files, every decision the
user makes can be expressed as a typed option up front, and the code path
runs unattended from start to finish. Differential expression, normalization,
peak calling, amplicon denoising, quantification import, batch correction and
annotation all fit.

Do **not** declare jobs for work whose value is interactive: visualization
browsers, iterative model tuning with a human in the loop, anything needing a
live R session mid-analysis. If some computation *inside* an interactive flow
is batch-shaped (fitting a model, scoring a matrix), consider exposing just
that.

One package can declare several jobs (one YAML plus script pair each, sharing
one name namespace), and a job need not expose everything a function can do:
expose the parameters that matter for batch use, hard-code sensible choices
for the rest, and keep the full flexibility in your R API.

Version 1.0 of the specification keeps the model small: single-file inputs
and outputs, five option types and one executor per job.

# Where to go next

- The
  [developer guide][guide]
  walks a maintainer from an empty directory to generated, validated
  wrappers, with the complete field reference, a release checklist, a
  troubleshooting table and an FAQ.
- The **DESeq2 worked example** in the BiocJobs repository
  ([`examples/DESeq2/`][deseq2])
  is the realistic counterpart to the toy job used here: 2 inputs, 3 outputs
  and 9 typed options exercising every option type, a defensively written
  analysis script, simulated test data, and all the generated artifacts under
  `generated/`, which CI validates with each consuming system's schema or
  linter (Galaxy XSD, GA4GH TES 1.1 schema, `miniwdl check`, `nextflow lint`,
  `kubeconform`). It is not installed with the package (and needs
  DESeq2), which is why this vignette uses the toy job instead.
- `help(package = "BiocJobs")` for the reference documentation of all
  exported functions.

# Session information

```{r session-info}
sessionInfo()
```

[guide]: https://github.com/almahmoud/BiocJobs/blob/main/docs/developer-guide.md
[deseq2]: https://github.com/almahmoud/BiocJobs/tree/main/examples/DESeq2
[tes]: https://github.com/ga4gh/task-execution-schemas
